---
title: "Exploratory Data Analysis Assignment 1"
author: "submitted by Vinay vernekar vrvernek@iu.edu"
date: "January 16, 2017"
output: pdf_document
---

```{r echo=FALSE}
library(ggplot2)
library(lattice)

setwd("C:\\dataset\\")

Tipdata<-read.table(file="tips.txt",header = TRUE)

#head(Tipdata,5)

Tipdata$tip_P<-(Tipdata$tip/Tipdata$total_bill)*100


```

__Q1.__ Does the percentage tipped follow a normal distribution? If not, how does the data differ from a normal distribution? Include ONE graph drawn with ggplot to support your answer.  

```{r echo=FALSE}
# Plotting QQ-Norm plot
ggplot(Tipdata, aes(sample=tip_P)) +stat_qq() +ggtitle("QQ Plot to check for Distribution") # fat tails observed
#ggplot(Tipdata, aes(x=tip_P)) + stat_ecdf()
ggplot(Tipdata, aes(sample=tip_P)) +stat_qq(distribution = qunif) # we notice curveture in the plot
#ggplot(Tipdata, aes(sample=tip_P)) +stat_qq(distribution = qunif)
#ggplot(Tipdata, aes(x=tip_P)) + geom_density()
```
__Observations:__

1) The data appears to be normally distributed, however the extreme value do appear to follow normal distribution.  
2) But when using uniform distribution it can be noted for a given of quantiles it follows an uniform distribution.    

__Q2.__ How does the distribution of tips change with party size? Include ONE plot to support your answer. (A set of faceted plots counts as one plot.)

```{r echo=FALSE}
ggplot(Tipdata, aes(sample = tip_P)) +stat_qq() +facet_wrap(~size, ncol=3)+ ggtitle("Tip Percent distribution for different size(attribute)")
#summary(as.factor(Tipdata$size))
```

__Observations:__  

1) We can notice that there are less points for size 1,5 and 6 and hence difficult to conclude that they are normally distributed. This also indicates that single and very big group sizes are rare.     
2) To a greater extent the distribution of tip percentage for size of 2,3 and 4 appear normal. But we can see that are still some observations that do not appear normally distributed and seem to be above or below the QQ line, these are outliers.  


__Q3.__ Using lm(), we can find that a linear model to predict percentage tipped from party size is
Percent tipped = 18.4 - 0.92 * size
Does it look like the model fits the data well? Include ONE plot of the residuals of this model to support your answer. (Again, a set of faceted plots counts as one plot.)	

```{r echo=FALSE}
percent_tippedlm = lm(tip_P ~ size, data=Tipdata)

tip_model_data=data.frame(Tipdata$total_bill,Tipdata$size,resi=residuals(percent_tippedlm))
#head(tip_model_data)

# plotting the data
ggplot(tip_model_data, aes(sample = resi)) +stat_qq()+ facet_wrap(~Tipdata.size, ncol=3)+ggtitle("Residual Plot Examination")
```
__Observations:__

1) Since there are a few data points for size 1,5,6 its hard to say that are really normally distributed.  

2) However the plots for the other size like 2,3 and 4 appear to be normally distributed with certain exemptions to extreme values or observations  




